### 2025-01-23 paper/icml25/first/sects/discussion.tex
minors

This advantage does not hold in the Dispersion scenario, where complexity increases substantially with additional agents. Nevertheless, we observe the emergence of cooperation across all scenarios, suggesting potential scalability to larger agent populations. The main bottleneck lies in the transformer architecture for the action world model, where the sequence length scales with the number of agents (sequence length = num. agents times num. train steps). With significantly large populations (~50 agents), the attention mechanism becomes computationally expensive, leading to prohibitively long memory requirements. In these scenarios, a linear attention mechanism (such as \cite{Beltagy20}) may be employed.

>>>>

This advantage does not hold in the Dispersion scenario, where complexity increases substantially with additional agents. Nevertheless, we observe the emergence of cooperation across all scenarios, suggesting potential scalability to larger agent populations. The main bottleneck lies in the transformer architecture for the action world model, where the sequence length scales with the number of agents (sequence length = num. agents times num. train steps). With significantly large populations (~50 agents), the attention mechanism becomes computationally expensive, leading to prohibitively large memory requirements. In these scenarios, a linear attention mechanism (such as \cite{Beltagy20}) may be employed.

### 2025-01-23 paper/icml25/first/sects/shacpp.tex
minors

As mentioned previously, we aim to address the limitations of SHAC by approximating the gradients of a differentiable environment ($R$ and $F$) with neural networks ($R_\theta$ and $F_\theta$). In particular, we aim to train a neural network to approximate the reward function by minimizing: 
\begin{align*}
    \underset{\tau\sim\pi}{E}\left[\hspace{-.7cm}\sum_{\hspace{.7cm}\scriptscriptstyle s_i,a_{ij},s_{i+1}\in\tau} \hspace{-.8cm}\left\vert\left\vert {\rewards{i}}_\theta(s_i, a_{ij}, s_{i+1}) -\rewards{i}(s_i, a_{ij}, s_{i+1}) \right\vert\right\vert^2\right] 
\end{align*}

In practice, this is achieved by leveraging a \emph{replay buffer} to cache the state-action-reward triplets obtained from the environment. The neural network is then trained to minimize the mean squared error (MSE) loss. Specifically, we minimize the loss $\mathcal{L}_\theta^{\rewards{i}}:(\states\times\actions{i}\times\states)^N\rightarrow\mathbb{R}$:
$$ \mathcal{L}_\theta^{\rewards{i}}(\batch) = \frac{1}{\bs}\sum_{s,a,s',r \in \batch}\left|\left| {\rewards{i}}_\theta(s, a) - r \right|\right|^2 $$
Where $B$ represents a batch of state-action-state-reward tuples sampled uniformly from the cache.

Similarly, we aim to train a neural network to approximate the transition function. However, to avoid vanishing or exploding gradients, we employ a non-recursive approach similar to Action World Model~\cite{Ma24}. This corresponds to train a neural network that given a starting observation from $\mathcal{S}$ and a sequence of $\mh$ actions from $\alla^\mh$, predicts the next $\mh$ observations from $\states^\mh$. We aim to minimize:

\begin{align*}
    \underset{\tau\sim\pi}{E}\left[\hspace{-1.2cm}\sum_{\scriptscriptstyle \hspace{1cm}\substack{s_t,\dots,s_{t+h+1},\\\hspace{.3cm}a_{t*},\dots,a_{t+h*}\in\tau}}\hspace{-1.2cm} \left|\left| \transition_{\theta}(s_t, a_{t*},\dots,a_{\scriptscriptstyle i+h*}) - [s_t,\dots,s_{\scriptscriptstyle t+h+1}] \right|\right|^2\right] 
\end{align*}

Where $a_{t*}=\{a_{ti}\}_{i\in\agents}$. In practice, this is achieved by caching trajectories and then training the neural network to minimize the MSE loss. That is, we minimize the loss $\mathcal{L}_\theta^{F}:\mathcal{T}^N\rightarrow\mathbb{R}$:
$$ \frac{1}{\bs\mh}\sum_{\tau \in B}\left|\left| \transition_{\theta}(s_0^\tau, a_{1*}^\tau, \dots, a_{\mh*}) - \left[ s^\tau_1, \dots, s^\tau_\mh \right] \right|\right|^2 $$
Where $B$ represents a batch of trajectories of length $\mh$ sampled uniformly from the cache.

>>>>

As mentioned previously, we aim to address the limitations of SHAC by approximating the gradients of a differentiable environment ($R$ and $F$) with neural networks ($R_\theta$ and $F_\theta$). In particular, we aim to train a neural network to approximate the reward function by minimizing: 
\begin{align*}
    \underset{\tau\sim\pi}{E}\left[\hspace{-.7cm}\sum_{\hspace{.7cm}\scriptscriptstyle s_i,a_{ij},s_{i+1}\in\tau} \hspace{-.8cm}\left( {\rewards{i}}_\theta(s_i, a_{ij}, s_{i+1}) -\rewards{i}(s_i, a_{ij}, s_{i+1}) \right)^2\right] 
\end{align*}

In practice, this is achieved by leveraging a \emph{replay buffer} to cache the state-action-reward triplets obtained from the environment. The neural network is then trained to minimize the mean squared error (MSE) loss. Specifically, we minimize the loss $\mathcal{L}_\theta^{\rewards{i}}:(\states\times\actions{i}\times\states)^N\rightarrow\mathbb{R}$:
$$ \mathcal{L}_\theta^{\rewards{i}}(\batch) = \frac{1}{\bs}\sum_{s,a,s',r \in \batch}\left( {\rewards{i}}_\theta(s, a) - r \right)^2 $$
Where $B$ represents a batch of state-action-state-reward tuples sampled uniformly from the cache.

Similarly, we aim to train a neural network to approximate the transition function. However, to avoid vanishing or exploding gradients, we employ a non-recursive approach similar to Action World Model~\cite{Ma24}. This corresponds to train a neural network that given a starting observation from $\mathcal{S}$ and a sequence of $\mh$ actions from $\alla^\mh$, predicts the next $\mh$ observations from $\states^\mh$. We aim to minimize:
\begin{align*}
    \underset{\tau\sim\pi}{E}\left[\hspace{-1.2cm}\sum_{\scriptscriptstyle \hspace{1cm}\substack{s_t,\dots,s_{t+h+1},\\\hspace{.3cm}a_{t*},\dots,a_{t+h*}\in\tau}}\hspace{-1.2cm} \left\vert\left\vert \transition_{\theta}(s_t, a_{t*},\dots,a_{\scriptscriptstyle i+h*}) - [s_t,\dots,s_{\scriptscriptstyle t+h+1}] \right\vert\right\vert^2\right] 
\end{align*}
Where $a_{t*}=\{a_{ti}\}_{i\in\agents}$. In practice, this is achieved by caching trajectories and then training the neural network to minimize the MSE loss. That is, we minimize the loss $\mathcal{L}_\theta^{F}:\mathcal{T}^N\rightarrow\mathbb{R}$:
$$ \frac{1}{\bs\mh}\sum_{\tau \in B}\left|\left| \transition_{\theta}(s_0^\tau, a_{1*}^\tau, \dots, a_{\mh*}) - \left[ s^\tau_1, \dots, s^\tau_\mh \right] \right|\right|^2 $$
Where $B$ represents a batch of trajectories of length $\mh$ sampled uniformly from the cache.
